Functional oxide layers for electrical isolation and chemical passivation of steel substrates

### 2022-08-11 Thesis/chapter/03-theoretical-background.tex
further stats: gisela Haertler

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Machine Learning and Statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Artificial Inteligence and Machine Learning}

The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at the Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
For a more in depth dive into the history of AI consult the beautifully written review\cite{Apter1982} of McCorduck's 1982 book "Machines Who Think"\cite{McCorduck1982}, which focuses also on the great minds behind advances of \gls{ai}.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{Howard2019}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test ingenioused by Alan Turing, is that a human communicates with an unknown entity (in written form) and must judge if they are dealing with a machine or a human being. 
It is easy for humans to come up with question to detect an AI, 
but when reading AI written articles\cite{gpt2020}, it's easy to see this test being passed in the near future.

But fear not, that doesn't mean that computers are sentient or more intelligent than humans\cite{searle1980,searle1999married} and certainly not that research is over. 
AI is still a young field, which is strongly growing and is gaining ubiquitous status. 
It is slowly creeping into every aspect of modern human life just like electricity around one hundred years ago. 
%Now, we can't - it's hard to - imagine to live without electricity. 
Realms in which AI is gaining traction are: 
%
playing board games (and beating humans)\cite{Silver2017,Feng1999,Campbell2002}, 
image recognition (in medicine)\cite{Li2020,Deo2015,Topol2019,Fujiyoshi2019}, 
in chemistry\cite{Westermayr2019,goh2017chemception,jha2018elemnet}, 
cyber security\cite{Sarker2021},
facial recognition (to prevent theft of toilet paper \cite{Andrews2017}),
financial sector (as robo-advisors)\cite{Littman2021},
natural language processing (NLP)\cite{Koroteev2021,Liu2021gpt,Parviainen2021} (which can also create pictures through scalable vector graphics (SVG) and code)
and even creative tasks like 
creating non existing faces\cite{Mansourifar2020}, 
create artwork (DALL-E 2)\cite{Marcus2022} or 
making video games\cite{Guzdial2016}.
%
It is nearly hard to find a field where \gls{ai} isn't used in some way. 
This steady incorporation of \gls{ai} leads to the so called \gls{ai} effect\cite{McCorduck1982,ai100}: 
certain fields get incorporated into \gls{ai} research and practice and after some time of general use are no more considered \gls{ai} (e.g. spam filter or web searches).
Google CEO Sundar Pichai even goes as far and said: 
"AI is one of the most important things humanity is working on. It is more profound than [...] electricity or fire"\cite{Hassan2020}

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous) versus classification (discrete). 
Independent from these $2\times2$ categories there are multiple ways to let machines learn from data.
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
%
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and in quantum mechanics too\cite{westermayr2020combining}.
Other common \td{architectures} are linear regression, kernel ridge regression, support vector regression.
There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Particle Swarm Optimization}
it has been shown that \gls{pso} works well on problems wher \gls{ga}s work well\cite{Kennedy1995}. 
It can be seen as in the middle of \gls{ga} (taking eons) and \gls{nn} taking seconds regarding time scale\cite{Kennedy1995}.
it is a stochastic process
are there iterative optimisation processes without stochastic? 
%
\subsubsection{things I should try if they work:}
\begin{itemize}
    \item GA 
    \item what is reinforcement learning? 
    \item PCA
    \item linear regression
    \item stat vs ML \url{https://medium.com/source-institute/ai-vs-statistics-c2485f9df126} and 
    \item https://towardsdatascience.com/no-machine-learning-is-not-just-glorified-statistics-26d3952234e3?gi=3f94b919de45
    \item https://towardsdatascience.com/are-you-aware-how-difficult-your-regression-problem-is-b7dae830652b calculate error/smoothness etc
\end{itemize}

@@

DOEs: randomization, latin square, orthogonal experiment design, full/2-lvl factorial design, placket burman, 
\url{https://www.jstor.org/stable/2238570}\cite{karlin1966optimal}
\url{https://rss.onlinelibrary.wiley.com/doi/abs/10.1111/j.2517-6161.1973.tb00944.x}\cite{whittle1973some}
maybe \url{https://www.sciencedirect.com/science/article/pii/S0167715212003343?via\%3Dihub}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Analysis of Variance} % ANOVA
The statistical significance of the experiment is determined by a ratio of two variances. This ratio is independent of several possible alterations to the experimental observations: Adding a constant to all observations does not alter significance. Multiplying all observations by a constant does not alter significance. 
So ANOVA statistical significance result is independent of constant bias and scaling errors as well as the units used in expressing observations. %https://en.wikipedia.org/wiki/Analysis_of_variance#Summary_of_assumptions
\td{make analysis with EMMA dataset and EMMA+DOE}
\url{https://www.scribbr.com/statistics/anova-in-r/} just make input categorical
\url{https://pypi.org/project/pynova/}
\textbf{Assumptions of ANOVA}
from \url{https://statsandr.com/blog/anova-in-r/}
\begin{itemize}
    \item variable type: continuous dependent variable and categorial qualitative independent variable (also called treatments or levels). The input variables are discretized an can therefore be seen as categorial.
    \item Independence: groups and levels should be independent
    \item Normality: the residual (that is the durchschnittliche dependance of uncontrollable variables ,like room temperature or humidity) should approximately follow a normal distribution. 
    \item Equality of variance: variance in differenct groups should be equal (compare with homoscedasticity).
    \item No outliers
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Linear Regression}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{AI vs statistics}
\begin{itemize}
	\item \td{let GPT-3 write}
	\item check out links at \url{https://yewtu.be/watch?v=PqbB07n\_uQ4}
	\item see pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
	\item 
		\href
		{https://doi.org/10.1177\%2F1352458520978648}
		{Machine and deep learning in MS research are just powerful statistics – No}
\end{itemize}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ds{
introduction 
    why I chose this practicum, 
        bcs wanted to check if lab is for me 
    why i chose to include ml 
        bcs I wanted what i learned during undergrad courses help me find a good solution and 
        wanted to dive deeper into the ml field as it is very interesting for 
what is ML? 
what kind of ML techniques? 
statistics:
what is the difference between ml and stat? 
    superficially very similar but different goals and diff rahmenbedingungen 
DOE is a very old, much information less experiments 
    how to design experiment: already many insights for next time 
statistics: 
    help me find insight from data 
    ANOVA 
    linear regression 
}

\subsection{Machine Learning and Statistics}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Why introduce machine learning to this project? 
I wanted to apply and delve into what I've been studying during my undergrad courses. 
This seemed like the perfect oppurtunity.

\subsubsection{Artificial Inteligence and Machine Learning}
%How can \gls{ai} and \gls{ml} be defined? 
The history of \gls{ai} goes back to the middle of the \nth{20} century. 
Researchers from the emerging field came together at the Dartmouth conference and the term "\gls{ai}" was coined\cite{McCarthy1955}. 
\Gls{ai}'s  history is beautifully depicted by McCorducks' 1982 book "Machines Who Think"\cite{McCorduck1982,Apter1982}, which focuses on the great minds behind the advances.
%For a more in depth dive into the history of AI consult the beautifully written review\cite{Apter1982} of McCorduck's 1982 book "Machines Who Think"\cite{McCorduck1982}, which focuses also on the great minds behind advances of \gls{ai}.
Pioneers like Alan Turing thought a lot about how to define, test and implement \gls{ai}\cite{Howard2019}. 
One example how to measure \gls{ai} is to let it play chess against a human\cite{Silver2017} 
(in 1997 a chess computer called Deep Blue won against the World Chess Champion Garry Kasparov for the first time\cite{Feng1999}).
Another test ingenioused by Alan Turing, is that a human communicates with an unknown entity (in written form) and must judge if they are dealing with a machine or a human being. 
It is easy for humans to come up with question to detect an AI, 
but when reading AI written articles\cite{gpt2020}, it's easy to see this test being passed in the near future.

But fear not, that doesn't mean that computers are sentient or more intelligent than humans\cite{searle1980,searle1999married} and certainly not that research is over. 
AI is still a young field, which is strongly growing and is gaining ubiquitous status. 
It is slowly creeping into every aspect of modern human life just like electricity around one hundred years ago. 
%Now, we can't - it's hard to - imagine to live without electricity. 
Realms in which AI is gaining traction are: 
%
playing board games (and beating humans)\cite{Silver2017,Feng1999,Campbell2002}, 
image recognition (in medicine)\cite{Li2020,Deo2015,Topol2019,Fujiyoshi2019}, 
in chemistry\cite{Westermayr2019,goh2017chemception,jha2018elemnet}, 
cyber security\cite{Sarker2021},
facial recognition (to prevent theft of toilet paper \cite{Andrews2017}),
financial sector (as robo-advisors)\cite{Littman2021},
natural language processing (NLP)\cite{Koroteev2021,Liu2021gpt,Parviainen2021} (which can also create pictures through scalable vector graphics (SVG) and code)
and even creative tasks like 
creating non existing faces\cite{Mansourifar2020}, 
create graphic artwork (DALL-E 2)\cite{Marcus2022} or 
making video games\cite{Guzdial2016}.
%
It is nearly hard to find a field where \gls{ai} isn't used in some way. 
This steady incorporation of \gls{ai} leads to the so called \gls{ai} effect\cite{McCorduck1982,ai100}: 
certain fields get incorporated into \gls{ai} research and practice and after some time of general use are no more considered \gls{ai} (e.g. spam filter or web searches).
Google CEO Sundar Pichai even goes as far and said: 
"AI is one of the most important things humanity is working on. It is more profound than [...] electricity or fire"\cite{Hassan2020}

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
%
\Gls{ml} is at the base of most \gls{ai}s.
It is a umbrella term for programs with instructions to learn from data, i.e. gain knowledge, categorise, predict and make decisions based on data. 
%
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous) versus classification (discrete). 
Independent from these $2\times2$ categories there are multiple ways to let machines learn from data.
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
%
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and in quantum mechanics too\cite{westermayr2020combining}.
Other common \td{architectures} are linear regression, kernel ridge regression, support vector regression.
There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
\td{This is exactly what I needed because every experiment is very time intensive.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Particle Swarm Optimization}

it has been shown that \gls{pso} works well on problems wher \gls{ga}s work well\cite{Kennedy1995}. 
It can be seen as in the middle of \gls{ga} (taking eons) and \gls{nn} taking seconds regarding time scale\cite{Kennedy1995}.
it is a stochastic process
are there iterative optimisation processes without stochastic? 
%

@@

DOEs: randomization, latin square, orthogonal experiment design, full/2-lvl factorial design, placket burman, 
\url{https://www.jstor.org/stable/2238570}\cite{karlin1966optimal}
\url{https://rss.onlinelibrary.wiley.com/doi/abs/10.1111/j.2517-6161.1973.tb00944.x}\cite{whittle1973some}
maybe \url{https://www.sciencedirect.com/science/article/pii/S0167715212003343?via\%3Dihub}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Analysis of Variance} % ANOVA
The statistical significance of the experiment is determined by a ratio of two variances. 
This ratio is independent of several possible alterations to the experimental observations: 
Adding a constant to all observations does not alter significance. 
Multiplying all observations by a constant does not alter significance. 
So ANOVA statistical significance result is independent of constant bias and scaling errors 
as well as the units used in expressing observations. %https://en.wikipedia.org/wiki/Analysis_of_variance#Summary_of_assumptions
\td{make analysis with EMMA dataset and EMMA+DOE}
\url{https://www.scribbr.com/statistics/anova-in-r/} just make input categorical
\url{https://pypi.org/project/pynova/}
\textbf{Assumptions of ANOVA}
from \url{https://statsandr.com/blog/anova-in-r/}
\begin{itemize}
    \item variable type: continuous dependent variable and categorial qualitative independent variable (also called treatments or levels). The input variables are discretized an can therefore be seen as categorial.
    \item Independence: groups and levels should be independent
    \item Normality: the residual (that is the durchschnittliche dependance of uncontrollable variables ,like room temperature or humidity) should approximately follow a normal distribution. 
    \item Equality of variance: variance in differenct groups should be equal (compare with homoscedasticity).
    \item No outliers
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Linear Regression}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{AI vs statistics}
As \gls{ml} failed me or rather I failed in \gls{ml} (see Learning from Data) I tried to extract as much insight as possible using statistics. 
What is statistics, though and what seperates it from \gls{ml}? 
\begin{itemize}
	\item \td{let GPT-3 write}
	\item check out links at \url{https://yewtu.be/watch?v=PqbB07n\_uQ4}
	\item see pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
	\item 
		\href
		{https://doi.org/10.1177\%2F1352458520978648}
		{Machine and deep learning in MS research are just powerful statistics – No}
\end{itemize}

@@

\subsubsection{things I should try if they work:}
\begin{itemize}
    \item GA 
    \item what is reinforcement learning? 
    \item PCA
    \item linear regression
    \item stat vs ML \url{https://medium.com/source-institute/ai-vs-statistics-c2485f9df126} and 
    \item https://towardsdatascience.com/no-machine-learning-is-not-just-glorified-statistics-26d3952234e3?gi=3f94b919de45
    \item https://towardsdatascience.com/are-you-aware-how-difficult-your-regression-problem-is-b7dae830652b calculate error/smoothness etc
\end{itemize}
